---
title: "Plot Gallery"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plot Gallery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse   = TRUE,
  comment    = "#>",
  fig.width  = 7,
  fig.height = 4
)
```

## Setup

All examples use the bundled synthetic `mv` object:

```{r load}
library(modelVis)

mv <- readRDS(
  file = system.file(
    "extdata", "example_mv.rds",
    package = "modelVis"
  )
)
summary(mv)
```

## Time-series trends

### Single trend with CI ribbon

`mv_plot_trend()` draws a line with an optional
confidence interval ribbon when `lwr` and `upr`
columns are present.

```{r trend-sb, fig.cap = "Spawning biomass with 95\\% CI ribbon."}
mv_plot_trend(
  df    = mv$spawning_biomass,
  ylab  = "Spawning biomass (kt)",
  title = "Spawning Biomass"
)
```

### Customising colour and reference lines

```{r trend-rec, fig.cap = "Recruitment with a reference line at R0."}
mv_plot_trend(
  df        = mv$recruitment,
  ylab      = "Recruitment (thousands)",
  title     = "Recruitment",
  colour    = "#2ca02c",
  ref_line  = mv$sr_params$R0,
  ref_label = "R0"
)
```

### Harvest rate (no CI)

When `lwr`/`upr` are `NA`, the ribbon is suppressed
automatically:

```{r trend-hr, fig.cap = "Total harvest rate time series."}
mv_plot_trend(
  df     = mv$legal_hr,
  ylab   = "Harvest rate",
  title  = "Legal Harvest Rate",
  colour = "#d62728"
)
```

### Log/linear toggle

Set `log_buttons = TRUE` to add a toggle between
linear and log y-axis scales:

```{r trend-log, fig.cap = "Spawning biomass with log/linear toggle buttons."}
mv_plot_trend(
  df          = mv$spawning_biomass,
  ylab        = "Spawning biomass (kt)",
  title       = "Log/Linear Toggle Demo",
  log_buttons = TRUE
)
```

### Multiple grouped trends

`mv_plot_trend_multi()` overlays lines for each level
of a grouping column:

```{r trend-multi, fig.cap = "Index predictions by fleet."}
idx_trend <- mv$index_fits[
  , c("year", "fleet", "pred"), drop = FALSE
]
names(idx_trend)[names(idx_trend) == "pred"] <- "est"

mv_plot_trend_multi(
  df    = idx_trend,
  group = "fleet",
  ylab  = "Predicted index",
  title = "Predicted Indices by Fleet"
)
```

## Observed vs predicted

### Single panel

`mv_plot_obs_pred()` shows observed markers and a
predicted line:

```{r obs-pred, fig.cap = "Survey index fit."}
surv <- mv$index_fits[
  mv$index_fits$fleet == "survey1",
  ,
  drop = FALSE
]
mv_plot_obs_pred(
  df    = surv,
  ylab  = "Survey index",
  title = "Survey 1 Index Fit"
)
```

### Faceted by fleet

`mv_plot_obs_pred_facet()` creates a vertically stacked
subplot for each fleet:

```{r obs-pred-facet, fig.height = 8, fig.cap = "Index fits faceted by fleet."}
mv_plot_obs_pred_facet(
  df    = mv$index_fits,
  facet = "fleet",
  ylab  = "Index"
)
```

## Age composition fits

### Grid of annual fits

`mv_plot_comp_grid()` shows observed (grey bars) and
predicted (coloured line + points) proportions at age
for the most recent years:

```{r comp-grid, fig.height = 8, fig.cap = "Age composition fits for fishery 1 (most recent 16 years)."}
f1_comps <- mv$age_comp_fits[
  mv$age_comp_fits$fleet == "fishery1",
  ,
  drop = FALSE
]
mv_plot_comp_grid(
  df             = f1_comps,
  bin_col        = "age",
  years_per_page = 16,
  ncol           = 4
)
```

### Average composition

`mv_plot_comp_avg()` averages observed and predicted
compositions across all years:

```{r comp-avg, fig.cap = "Time-averaged age composition for fishery 1."}
mv_plot_comp_avg(
  df      = f1_comps,
  bin_col = "age",
  ylab    = "Mean proportion",
  title   = "Fishery 1 -- Average Age Composition"
)
```

### Single year

`mv_plot_comp_fit()` shows a single year:

```{r comp-single, fig.cap = "Age composition fit for fishery 1 in 2020."}
yr2020 <- f1_comps[
  f1_comps$year == 2020,
  ,
  drop = FALSE
]
mv_plot_comp_fit(
  df      = yr2020,
  bin_col = "age",
  title   = "Fishery 1 -- 2020"
)
```

## Residual diagnostics

### Bubble plot

`mv_plot_bubble()` shows residuals as bubbles sized by
absolute value and coloured by sign (red = positive,
blue = negative):

```{r bubble, fig.cap = "Age composition residual bubble plot for survey 1."}
surv_comps <- mv$age_comp_fits[
  mv$age_comp_fits$fleet == "survey1",
  ,
  drop = FALSE
]
surv_comps$residual <- log(surv_comps$obs) -
  log(surv_comps$pred)
mv_plot_bubble(
  df    = surv_comps,
  x_col = "year",
  y_col = "age",
  title = "Survey 1 -- Composition Residuals"
)
```

### Four-panel residual diagnostics

`mv_plot_resid_panel()` creates a 2$\times$2 subplot
with: residuals vs year, residuals vs predicted, QQ
plot, and histogram:

```{r resid-panel, fig.height = 6, fig.cap = "Index residual diagnostics for survey 1."}
surv_idx <- mv$index_fits[
  mv$index_fits$fleet == "survey1",
  ,
  drop = FALSE
]
mv_plot_resid_panel(
  df    = surv_idx,
  title = "Survey 1 -- Index Residuals"
)
```

## Selectivity at age

`mv_plot_selectivity()` overlays selectivity curves for
each fleet:

```{r selectivity, fig.cap = "Estimated selectivity-at-age by fleet."}
mv_plot_selectivity(
  df    = mv$selectivity_at_age,
  ylab  = "Selectivity",
  title = "Selectivity at Age"
)
```

For sex-structured models, set `by_sex = TRUE` to
create side-by-side panels.

## Stock-recruitment

`mv_plot_sr()` creates a scatter of recruitment vs
spawning biomass, with points coloured by year:

```{r sr, fig.cap = "Stock-recruitment relationship."}
mv_plot_sr(
  df    = mv$stock_recruit,
  ylab  = "Recruitment (thousands)",
  xlab  = "Spawning biomass (kt)",
  title = "Stock-Recruitment"
)
```

Set `show_labels = TRUE` to annotate every 5th year.

## Kobe phase plot

`mv_plot_kobe()` shows stock status relative to
reference points in a four-quadrant phase plot:

```{r kobe, fig.cap = "Kobe phase plot showing stock trajectory."}
b_ratio <- mv$spawning_biomass$est /
  mv$ref_points$SBmsy
f_ratio <- mv$legal_hr$est /
  mv$ref_points$Fmsy

mv_plot_kobe(
  biomass   = b_ratio,
  mortality = f_ratio,
  years     = mv$meta$years,
  xlab      = "B/Bmsy",
  ylab      = "U/Umsy",
  title     = "Kobe Phase Plot"
)
```

The terminal year is highlighted with a red star.

## Equilibrium reference curves

`mv_plot_refcurves()` shows equilibrium yield and
biomass as a function of harvest rate, with reference
point annotations:

```{r refcurves, fig.height = 6, fig.cap = "Equilibrium yield and biomass curves."}
mv_plot_refcurves(
  ref_curves = mv$ref_curves,
  ref_points = mv$ref_points,
  title      = "Equilibrium Reference Curves"
)
```

## Age $\times$ year heatmap

`mv_plot_surface()` renders an age-by-year heatmap:

```{r surface, fig.cap = "Predicted age composition heatmap for fishery 1."}
surf_df <- f1_comps[
  , c("year", "age", "pred"), drop = FALSE
]
names(surf_df)[names(surf_df) == "pred"] <- "est"
mv_plot_surface(
  df    = surf_df,
  ylab  = "Age",
  title = "Fishery 1 -- Predicted Composition"
)
```

## Model comparison

`mv_plot_compare()` overlays a time-series component
from multiple `mv` objects. Each object's `meta$label`
is used as the legend entry.

```{r compare, fig.cap = "Spawning biomass comparison between two scenarios."}
# Create a second mv with shifted biomass
mv2_sb <- mv$spawning_biomass
mv2_sb$est <- mv2_sb$est * 1.15
mv2_sb$lwr <- mv2_sb$lwr * 1.15
mv2_sb$upr <- mv2_sb$upr * 1.15

mv2 <- new_mv(
  meta = modifyList(
    x   = mv$meta,
    val = list(label = "High productivity")
  ),
  spawning_biomass = mv2_sb,
  recruitment      = mv$recruitment
)

mv_plot_compare(
  mv_list   = list(mv, mv2),
  component = "spawning_biomass",
  ylab      = "Spawning biomass (kt)",
  title     = "Scenario Comparison"
)
```

## Theming and layout

All plots use `mv_layout()` for consistent styling.
You can modify the returned plotly object directly:

```{r custom-layout}
p <- mv_plot_trend(
  df   = mv$spawning_biomass,
  ylab = "SB (kt)"
)

# Override with plotly::layout()
p <- plotly::layout(
  p = p,
  title = list(
    text = "Custom Title",
    font = list(size = 18, color = "#333")
  ),
  paper_bgcolor = "#f8f8f8"
)
p
```

Use `mv_config()` to customise the plotly modebar
(the toolbar that appears on hover).
