---
title: "Writing an Extractor"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Writing an Extractor}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

## Overview

To connect a new stock assessment model to modelVis,
you write a single S3 method: `extract.mymodel()`.
This method lives in **your model package** (not in
modelVis) and converts your model's output into a
standardised `mv` object.

The workflow:

1. Define `extract.mymodel()` in your model package
2. Register the S3 method in your NAMESPACE
3. Optionally register a fingerprint for auto-detection
4. Users call `mv_dashboard()` or `extract()` --- S3
   dispatch routes to your method

## Step 1: Define the method

Your method must accept `fit` (the model output) and
`label` (a character string) and return an `mv` object
via `new_mv()`.

Here is a minimal skeleton:

```{r skeleton, eval = FALSE}
#' Extract mv object from MyModel
#'
#' @param fit MyModel output list
#' @param label Character label for this run
#' @param ... Additional arguments (unused)
#' @return An mv object
#' @export
extract.mymodel <- function(fit, label = "", ...) {

  # 1. Pull out dimensions and names
  years      <- fit$years
  ages       <- fit$ages
  fleetNames <- fit$fleet_names

  # 2. Build metadata
  meta <- list(
    model_type       = "mymodel",
    label            = label,
    years            = years,
    ages             = ages,
    sex_names        = "Combined",
    fleet_names      = fleetNames,
    area_names       = "All",
    has_length_comps = FALSE,
    has_discards     = FALSE,
    has_mcmc         = FALSE,
    fleet_type       = fit$fleet_type
  )

  # 3. Reshape data into standard data.frames
  spawning_biomass <- data.frame(
    year = years,
    est  = fit$ssb
  )

  recruitment <- data.frame(
    year = years,
    est  = fit$rec
  )

  # 4. Assemble
  modelVis::new_mv(
    meta             = meta,
    spawning_biomass = spawning_biomass,
    recruitment      = recruitment
  )
}
```

## Step 2: Component conventions

Each data component has required columns. NULL
components are permitted and cause their dashboard
sections to be skipped.

### Time-series components

Used for: `spawning_biomass`, `recruitment`,
`rec_devs`, `legal_hr`, `total_biomass`,
`legal_biomass`, `sublegal_biomass`.

| Column | Required | Description           |
|--------|----------|-----------------------|
| `year` | yes      | Integer year          |
| `est`  | yes      | Point estimate        |
| `lwr`  | no       | Lower CI bound        |
| `upr`  | no       | Upper CI bound        |

Use `mv_ensure_ci()` to add `NA` CI columns if your
model does not produce uncertainty intervals:

```{r ensure-ci, eval = FALSE}
spawning_biomass <- modelVis::mv_ensure_ci(
  df = spawning_biomass
)
```

### Fit components

Used for: `index_fits`, `catch_fits`, `discard_fits`.

| Column     | Required | Description            |
|------------|----------|------------------------|
| `year`     | yes      | Integer year           |
| `obs`      | yes      | Observed value         |
| `pred`     | yes      | Predicted value        |
| `fleet`    | no       | Fleet name             |
| `residual` | no       | Log residual           |

### Composition fits

Used for: `age_comp_fits`, `len_comp_fits`.

| Column     | Required | Description            |
|------------|----------|------------------------|
| `age`/`length` | yes | Bin value              |
| `year`     | yes      | Integer year           |
| `obs`      | yes      | Observed proportion    |
| `pred`     | yes      | Predicted proportion   |
| `fleet`    | yes      | Fleet name             |
| `sex`      | no       | Sex label              |

### Selectivity

| Column  | Required | Description            |
|---------|----------|------------------------|
| `age`   | yes      | Age class              |
| `sex`   | no       | Sex label              |
| `est`   | yes      | Selectivity value      |
| `fleet` | yes      | Fleet name             |

### Reference points

A named list. Common elements:

```{r refpts, eval = FALSE}
ref_points <- list(
  Fmsy  = 0.18,
  SBmsy = 25.3,
  lUmsy = 0.16,
  lMSY  = 4.5,
  B0    = 55.0,
  R0    = 10.0
)
```

### Reference curves

A data.frame with columns `legalU`, `yield`, and
`biomass` for equilibrium yield/biomass as a function
of legal harvest rate.

### Stock-recruitment

| Column        | Required | Description        |
|---------------|----------|--------------------|
| `ssb`         | yes      | Spawning biomass   |
| `recruitment` | yes      | Recruitment        |
| `year`        | yes      | Integer year       |

## Step 3: Using tidy helpers

Model arrays rarely match the long-format data.frames
that plot functions expect. modelVis exports helpers
for common reshaping tasks.

### Vectors to time-series

```{r tidy-ts}
library(modelVis)

ssb_vec <- c(50, 48, 45, 43, 44, 46)

mv_tidy_ts(
  x         = ssb_vec,
  years     = 2000:2005,
  value_col = "est"
)
```

### Matrices to long format

```{r tidy-mat}
# Selectivity: ages (rows) x fleets (cols)
sel_mat <- matrix(
  data     = c(0.1, 0.5, 0.9, 1.0,
               0.2, 0.7, 1.0, 1.0),
  nrow     = 4,
  ncol     = 2,
  dimnames = list(1:4, c("trawl", "survey"))
)

mv_tidy_mat(
  x          = sel_mat,
  row_name   = "age",
  col_name   = "fleet",
  value_col  = "est"
)
```

### 3D arrays

```{r tidy-3d}
# Catch-at-age: ages x fleets x years
caa <- array(
  data     = runif(n = 24),
  dim      = c(4, 2, 3),
  dimnames = list(
    age   = 1:4,
    fleet = c("trawl", "survey"),
    year  = 2020:2022
  )
)

head(mv_tidy_3d(
  x          = caa,
  dim_names  = c("age", "fleet", "year"),
  value_col  = "est"
))
```

### MCMC posteriors to CI time-series

When MCMC samples are available, `mv_quantile_ts()`
computes the median and credible interval:

```{r quantile-ts}
# Fake posterior: 500 iterations x 6 years
post_mat <- matrix(
  data = rnorm(n = 3000, mean = 50, sd = 5),
  nrow = 500,
  ncol = 6
)

mv_quantile_ts(
  x     = post_mat,
  years = 2000:2005,
  probs = c(0.025, 0.5, 0.975)
)
```

## Step 4: Register the S3 method

In your model package's `R/extract_mv.R` (or similar),
use roxygen2 to export the method:

```{r roxygen, eval = FALSE}
#' @export
extract.mymodel <- function(fit, label = "", ...) {
 # ...
}
```

Then in your NAMESPACE (generated by roxygen2):

```
S3method(extract, mymodel)
```

Your package must list `modelVis` in `Suggests:` (not
`Imports:`), and guard the method body with
`requireNamespace()`:

```{r guard, eval = FALSE}
extract.mymodel <- function(fit, label = "", ...) {
  if (!requireNamespace("modelVis", quietly = TRUE))
    stop(
      "Package 'modelVis' is required. ",
      "Install from LFR-code/modelVis."
    )
  # ... build components ...
  modelVis::new_mv(meta = meta, ...)
}
```

## Step 5: Auto-detection (optional)

modelVis can auto-detect your model type from an
unclassed list by checking structural fingerprints.
To register a fingerprint, submit a pull request
adding a check to `.detect_model_type()` in
`R/extract.R`:
```{r fingerprint, eval = FALSE}
.detect_model_type <- function(fit) {
  if (!is.list(fit)) return(NULL)
  # Existing: sableOpMod
  if (!is.null(fit$repOpt) && !is.null(fit$data))
    return("sableOpMod")
  # New: mymodel
  if (!is.null(fit$mymodel_field))
    return("mymodel")
  NULL
}
```

With auto-detection, users do not need to manually set
the class before calling `extract()`.

## Worked example: SISCA

The SISCA stock assessment model registers
`extract.sisca()` in its own package. Key patterns:

### Looping over populations and gears

SISCA stores arrays indexed by population, gear, and
time. The extractor loops over these dimensions and
builds row-list data.frames:

```{r sisca-loop, eval = FALSE}
idx_rows <- list()
for (p in seq_len(nP))
  for (g in seq_len(nG)) {
    obs_vec  <- as.numeric(data$I_pgt[p, g, ])
    pred_vec <- as.numeric(
      rep$qhat_pg[p, g] * rep$vulnB_pgt[p, g, ]
    )
    obs_vec[obs_vec < 0] <- NA
    residual <- ifelse(
      test = !is.na(obs_vec) & obs_vec > 0 &
             pred_vec > 0,
      yes  = log(obs_vec) - log(pred_vec),
      no   = NA_real_
    )
    idx_rows[[length(idx_rows) + 1]] <- data.frame(
      year     = assessYrs,
      obs      = obs_vec,
      pred     = pred_vec,
      fleet    = fleetNames[g],
      area     = areaNames[p],
      residual = residual
    )
  }
index_fits <- do.call(
  what = rbind, args = idx_rows
)
```

### Deriving fleet type from the control file

Fleet type (fishery vs survey) is read from the
control file, not hardcoded:

```{r sisca-fleet, eval = FALSE}
fleetType <- rep(1L, nG)
fleetType[ctl$data$fleetType == 0] <- 0L
```

### Normalising age compositions

Observed and predicted proportions are normalised
before storage:

```{r sisca-comp, eval = FALSE}
obs_sum <- sum(obs_vec, na.rm = TRUE)
if (obs_sum > 0) obs_vec <- obs_vec / obs_sum
pred_sum <- sum(pred_vec, na.rm = TRUE)
if (pred_sum > 0) pred_vec <- pred_vec / pred_sum
```

### Conditional components

Components that may not exist (e.g., reference points
from an optional post-processing step) are set to NULL
rather than omitted:

```{r sisca-conditional, eval = FALSE}
ref_points <- NULL
if (!is.null(fit$refPts)) {
  rp <- fit$refPts
  ref_points <- list(
    Fmsy  = rp$FmsyRefPts$Fmsy_p[1],
    SBmsy = rp$FmsyRefPts$SBeqFmsy_p[1],
    lUmsy = rp$FmsyRefPts$Umsy_p[1],
    B0    = rep$B0_p[1],
    R0    = rep$R0_p[1]
  )
}
```

## Testing your extractor

A basic test suite should verify:

1. The method returns an `mv` object
2. Required metadata fields are present
3. Key components have the expected columns
4. The dashboard renders without error

```{r test, eval = FALSE}
test_that("extract.mymodel returns valid mv", {
  fit <- readRDS(file = "test_fit.rds")
  mv  <- extract(fit = fit, label = "test")

  expect_s3_class(object = mv, class = "mv")
  expect_equal(
    object   = mv$meta$model_type,
    expected = "mymodel"
  )
  expect_true(
    object = all(
      c("year", "est") %in%
        names(mv$spawning_biomass)
    )
  )
  expect_true(
    object = all(
      c("year", "obs", "pred") %in%
        names(mv$index_fits)
    )
  )
})
```
