---
title: "Introduction to modelVis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to modelVis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 4
)
```

## What is modelVis?

modelVis generates interactive HTML dashboards for
fisheries stock assessment model diagnostics. It
decouples visualisation from modelling by defining a
standard intermediate object (the `mv` object) that any
model package can produce.

The pipeline is:

```
model output --> extract() --> mv object
                                 |
                      +----------+---------+
                      |                    |
              mv_dashboard()      mv_plot_*() functions
              (full dashboard)   (individual plots)
```

Each model package writes one `extract.mymodel()` method
to convert its output into an `mv` object. modelVis
handles everything downstream.

## Installation

```{r install, eval = FALSE}
devtools::install_github("LFR-code/modelVis")
```

## The mv object

An `mv` object is an S3 list with class `"mv"`. It
contains:
- **meta** --- a required metadata list (model type,
  label, years, ages, sex names, fleet names, area
  names, and flags)
- **data components** --- named data.frames following
  specific column conventions (all optional; NULL
  components cause their dashboard sections to be
  skipped)

### Metadata fields

| Field              | Type      | Description           |
|--------------------|-----------|-----------------------|
| `model_type`       | character | Model class name      |
| `label`            | character | Run label             |
| `years`            | integer   | Assessment years      |
| `ages`             | integer   | Age classes           |
| `sex_names`        | character | Sex labels            |
| `fleet_names`      | character | Fleet labels          |
| `area_names`       | character | Area labels           |
| `has_length_comps` | logical   | Length data flag       |
| `has_discards`     | logical   | Discard data flag     |
| `has_mcmc`         | logical   | MCMC posterior flag   |
| `fleet_type`       | integer   | 1 = fishery, 0 = survey |

### Data component conventions

**Time-series** (spawning biomass, recruitment,
harvest rate, etc.): must have `year` and `est`
columns. Optional `lwr` and `upr` columns trigger
confidence interval ribbons.

**Fit components** (index fits, catch fits): must have
`year`, `obs`, and `pred` columns. Optional `fleet`
column for faceting and `residual` for diagnostics.

**Composition fits** (age, length): must have a bin
column (`age` or `length`), `year`, `obs`, `pred`,
and `fleet`.

**Reference points**: a named list with elements like
`Fmsy`, `SBmsy`, `lUmsy`, `B0`, `R0`.

### Creating an mv object

The `new_mv()` constructor validates required metadata
and column names:

```{r new-mv}
library(modelVis)

mv <- new_mv(
  meta = list(
    model_type  = "demo",
    label       = "Minimal example",
    years       = 2000:2005,
    ages        = 1:5,
    sex_names   = "Combined",
    fleet_names = "survey1",
    area_names  = "All"
  ),
  spawning_biomass = data.frame(
    year = 2000:2005,
    est  = c(50, 48, 45, 43, 44, 46)
  )
)

print(mv)
```

### Inspecting an mv object

```{r summary}
summary(mv)
```

## Loading the example mv object

modelVis ships with a synthetic example for use in
vignettes and testing:

```{r load-example}
example_mv <- readRDS(
  file = system.file(
    "extdata", "example_mv.rds",
    package = "modelVis"
  )
)
print(example_mv)
```

## Generating a dashboard

The `mv_dashboard()` function is the main entry point
for full dashboards. It takes a raw model fit object,
calls `extract()` internally, and renders an interactive
flexdashboard.

```{r dashboard, eval = FALSE}
# From a model fit object (calls extract() internally)
mv_dashboard(
  fit        = model_reports,
  label      = "Base case 2025",
  output_file = "diagnostics.html"
)

# Or pass NULL for output_file to open an interactive
# viewer in RStudio
mv_dashboard(fit = model_reports, label = "Base case")
```

For model comparison, pass additional fits via the
`comp` argument:

```{r compare, eval = FALSE}
mv_dashboard(
  fit   = base_reports,
  label = "Base case",
  comp  = list(
    "Low M"  = lowM_reports,
    "High h" = highH_reports
  ),
  output_file = "comparison.html"
)
```

## Using individual plot functions

Every `mv_plot_*()` function accepts standard
data.frames and returns a plotly object. You can use
them independently of the dashboard:
```{r individual-plot}
mv_plot_trend(
  df    = example_mv$spawning_biomass,
  ylab  = "Spawning biomass (kt)",
  title = "Spawning Biomass"
)
```

See `vignette("modelvis-plots")` for a full gallery.

## Writing an extractor for your model

To connect a new model to modelVis, you write a single
S3 method: `extract.mymodel()`. This method reshapes
your model output into `new_mv()` format.

See `vignette("modelvis-extractor")` for a step-by-step
guide.

## Colour palettes

modelVis provides consistent colour palettes:

- `mv_colours(n)` --- general-purpose palette (10
  colours, interpolates beyond 10)
- `mv_fleet_colours(fleet_names)` --- named vector
  keyed to fleet names
- `mv_sex_colours(sex_names)` --- standard sex colours
  (Male = blue, Female = red, Combined = grey)
- `mv_kobe_colours()` --- status zone quadrant colours

```{r colours}
mv_colours(n = 5)
mv_sex_colours(sex_names = c("Male", "Female"))
```

## Tidy helpers for extractors

modelVis exports helper functions for converting model
arrays into the long-format data.frames that plot
functions expect:

| Function         | Input               | Output         |
|------------------|---------------------|----------------|
| `mv_tidy_ts()`   | named vector        | year + est     |
| `mv_tidy_mat()`  | matrix              | row + col + est|
| `mv_tidy_3d()`   | 3D array            | long data.frame|
| `mv_tidy_4d()`   | 4D array            | long data.frame|
| `mv_tidy_5d()`   | 5D array            | long data.frame|
| `mv_quantile_ts()` | MCMC matrix       | year + est + CI|
| `mv_ensure_ci()` | data.frame          | adds lwr/upr   |
| `mv_year_seq()`  | fYear + nT          | integer vector |
