
Age-comp Correlations {data-navmenu="Diagnostics"}
=====================================

Row {.tabset}
-------------------------------------

### About

Estimated age-composition correlation matrices (`Corr_gaa`) from the
model. Each cell shows the correlation between age classes $a$ and $a'$
in the compositional likelihood for that fleet.
Upper triangle: coloured circles (area ∝ |r|, blue = positive,
red = negative). Lower triangle: numeric values. Matches the
Compositional Likelihood Correlation Matrices in the fit report.

```{r comp_corr_about}
cc <- mv$comp_corr
cat(sprintf("%d fleet(s) have correlation matrices.\n", length(cc)))
for (item in cc)
  cat(sprintf("  %s : %d x %d\n", item$fleet,
              nrow(item$matrix), ncol(item$matrix)))
```

### Correlation Matrices

```{r comp_corr_matrices, results='asis', eval=length(mv$comp_corr) > 0 && requireNamespace("corrplot", quietly = TRUE)}
cc   <- mv$comp_corr
nmat <- length(cc)
for (i in seq_len(nmat)) {
  tmp <- tempfile(fileext = ".png")
  grDevices::png(tmp, width = 600L, height = 620L,
                 units = "px", res = 96L)
  corrplot::corrplot.mixed(
    corr       = cc[[i]]$matrix,
    lower.col  = "black",
    number.cex = 0.7,
    mar        = c(0, 0, 2, 0))
  graphics::mtext(
    text = cc[[i]]$fleet,
    side = 3, line = 0.3, cex = 1.1, font = 2)
  grDevices::dev.off()
  sz  <- file.info(tmp)$size
  b64 <- jsonlite::base64_enc(readBin(tmp, "raw", sz))
  cat('<img src="data:image/png;base64,', b64,
      '" style="width:600px;max-width:100%;',
      'display:block;margin:0 auto 24px;" />\n', sep = "")
}
```
