
Age-comp Correlations {data-navmenu="Diagnostics"}
=====================================

Row {.tabset}
-------------------------------------

### About

Estimated age-composition correlation matrices (`Corr_gaa`) from the model.
Each cell shows the correlation between age classes $a$ and $a'$ in the
multinomial/Dirichlet-multinomial age-composition likelihood for that fleet.
Strong off-diagonal correlations indicate that the model is compensating for
systematic age-reading or selectivity patterns.

```{r comp_corr_about}
cc <- mv$comp_corr
cat(sprintf("%d fleet(s) have correlation matrices.\n", length(cc)))
for (item in cc)
  cat(sprintf("  %s : %d Ă— %d\n", item$fleet,
              nrow(item$matrix), ncol(item$matrix)))
```

### Heatmaps

```{r comp_corr_heatmaps}
cc <- mv$comp_corr
hm_plots <- lapply(cc, function(item) {
  mat   <- item$matrix
  n_age <- nrow(mat)
  lbl   <- item$fleet

  # Colour scale: blue-white-red centred at 0
  p <- plot_ly(
    z         = mat,
    x         = colnames(mat),
    y         = rownames(mat),
    type      = "heatmap",
    colorscale = list(
      c(0, "#2166ac"), c(0.5, "#f7f7f7"), c(1, "#d6604d")
    ),
    zmin = -1, zmax = 1,
    showscale  = TRUE,
    hovertemplate = paste0(
      "Age row %{y} Ă— col %{x}<br>r = %{z:.3f}<extra>", lbl, "</extra>"
    )
  )
  lay <- modelVis::mv_layout()
  lay$xaxis$title <- "Age"
  lay$yaxis$title <- "Age"
  lay$annotations <- list(mv_top_label(label = lbl))
  do.call(layout, c(list(p = p), lay))
})

if (length(hm_plots) > 1L) {
  n_col <- min(length(hm_plots), 3L)
  n_row <- ceiling(length(hm_plots) / n_col)
  p_out <- do.call(subplot, c(hm_plots, list(
    nrows = n_row, shareX = FALSE, shareY = FALSE,
    titleX = TRUE, titleY = TRUE, margin = 0.08
  )))
  modelVis::mv_config(p = layout(p = p_out, showlegend = FALSE))
} else if (length(hm_plots) == 1L) {
  modelVis::mv_config(p = hm_plots[[1L]])
}
```
