---
title: "`r if (exists('tmp')) tmp$mv$meta$label else 'Simulation Report'`"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    css: !expr system.file("rmd", "style.css", package = "modelVis")
    theme: cosmo
    includes:
      in_header: !expr system.file("rmd", "viewport.html", package = "modelVis")
---

```{r setup, include=FALSE}
library(plotly)
library(modelVis)
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.height = 6,
  fig.width  = 9
)

# Retrieve data passed from mv_sim_dashboard()
mv <- tmp$mv

# Convenience references
meta       <- mv$meta
years      <- meta$years
ages       <- meta$ages
fleets     <- meta$fleet_names
tMP        <- meta$tMP
fYear      <- meta$fYear
pT         <- meta$pT
nReps      <- meta$n_reps
tMP_year   <- fYear + tMP - 1
fleet_type <- if (!is.null(meta$fleet_type)) {
  meta$fleet_type
} else {
  rep(1L, length(fleets))
}
comm_fleets   <- fleets[fleet_type == 1]
survey_fleets <- fleets[fleet_type == 0]

# Flags
has_catch    <- !is.null(mv$catch) &&
                any(mv$catch$est > 0, na.rm = TRUE)
has_refpts   <- !is.null(mv$ref_points)
has_tac      <- !is.null(mv$tac)
has_indices  <- !is.null(mv$indices) &&
                nrow(mv$indices) > 0
has_sr       <- !is.null(mv$stock_recruit) &&
                nrow(mv$stock_recruit) > 0
has_hcr      <- !is.null(mv$hcr_pars)
is_nofish    <- isTRUE(meta$mp_name == "NoFish")
has_cond_valid <- !is.null(mv$cond_valid)

# Reference point values for annotation
Bmsy <- if (has_refpts) mv$ref_points$Bmsy else NULL
Fmsy <- if (has_refpts) mv$ref_points$Fmsy else NULL
B0   <- if (has_refpts) mv$ref_points$B0   else NULL

# All good reps for slider
good_reps <- meta$good_reps

# tMP vertical line shape (reusable)
tMP_vline <- list(
  type = "line",
  x0 = tMP_year, x1 = tMP_year,
  y0 = 0, y1 = 1, yref = "paper",
  line = list(color = "#666666", width = 1, dash = "dot")
)
```

# Overview

## Row {data-height=80}

### Simulation details

**Scenario:** `r meta$scenario` |
**MP:** `r meta$mp_name` |
**Species:** `r meta$species` |
**Stock:** `r meta$stock` |
**Replicates:** `r nReps` |
**Historical:** `r fYear`--`r tMP_year - 1` |
**Projection:** `r tMP_year`--`r max(years)`

```{r cond_valid_header, eval=has_cond_valid, results='asis'}
cv <- mv$cond_valid$max_re
# Catch uses time-averaged weights by design; exclude
# from the pass/fail check
check_vars <- cv$variable != "Catch"
warn_flag <- any(cv$max_abs_re[check_vars] > 0.05)
icon <- if (warn_flag) "\u26A0\uFE0F" else "\u2705"
cat(paste0(
  "**Conditioning validation:** ", icon, " "
))
for (i in seq_len(nrow(cv))) {
  re_pct <- round(cv$max_abs_re[i] * 100, 1)
  cat(paste0(cv$variable[i], ": ", re_pct, "% "))
}
cat("\n")
```

## Row

### Spawning biomass

```{r sb_envelope, fig.height=3}
p <- mv_plot_trend(
  df       = mv$spawning_biomass,
  ylab     = "Spawning biomass (t)",
  ref_line = Bmsy,
  ref_label = if (has_refpts) "Bmsy" else NULL
)
p <- plotly_build(p)
p$x$layout$shapes <- c(p$x$layout$shapes, list(tMP_vline))
p
```

## Row

### Total catch

```{r catch_envelope, eval=has_catch, fig.height=3}
p <- mv_plot_trend(
  df       = mv$catch,
  ylab     = "Catch (t)",
  colour   = "#d62728"
)
p <- plotly_build(p)
p$x$layout$shapes <- c(p$x$layout$shapes, list(tMP_vline))
p
```

```{r catch_placeholder, eval=!has_catch, fig.height=3}
plotly_empty() %>%
  layout(
    title = list(
      text = "No fishing in this scenario",
      font = list(size = 14, color = "#999999")
    )
  )
```

## Row

### Recruitment

```{r rec_envelope, fig.height=3}
p <- mv_plot_trend(
  df     = mv$recruitment,
  ylab   = "Recruitment (age 0)",
  colour = "#2ca02c"
)
p <- plotly_build(p)
p$x$layout$shapes <- c(p$x$layout$shapes, list(tMP_vline))
p
```

## Row

### Total biomass

```{r tb_envelope, fig.height=3}
p <- mv_plot_trend(
  df       = mv$total_biomass,
  ylab     = "Total biomass (t)",
  colour   = "#ff7f0e",
  ref_line = B0,
  ref_label = if (has_refpts) "B0" else NULL
)
p <- plotly_build(p)
p$x$layout$shapes <- c(p$x$layout$shapes, list(tMP_vline))
p
```

## Row

### Natural mortality (mean over ages)

```{r m_envelope, eval=!is.null(mv$nat_mort), fig.height=3}
p <- mv_plot_trend(
  df     = mv$nat_mort,
  ylab   = "M (mean over ages)",
  colour = "#8c564b"
)
p <- plotly_build(p)
p$x$layout$shapes <- c(p$x$layout$shapes, list(tMP_vline))
p
```


# Fishing {`r if (is_nofish) ".hidden"`}

## Row

### Harvest rate by fleet

```{r hr_fleet, eval=has_catch}
hr_comm <- mv$harvest_rate[
  mv$harvest_rate$fleet %in% comm_fleets, ,
  drop = FALSE
]
if (nrow(hr_comm) > 0) {
  p <- mv_plot_trend(
    df     = hr_comm,
    ylab   = "Apical F",
    colour = "#1f77b4",
    ref_line = Fmsy,
    ref_label = if (has_refpts) "Fmsy" else NULL
  )
  p <- plotly_build(p)
p$x$layout$shapes <- c(p$x$layout$shapes, list(tMP_vline))
  p
}
```

### Catch by fleet

```{r catch_fleet, eval=has_catch}
ct_comm <- mv$catch_by_fleet[
  mv$catch_by_fleet$fleet %in% comm_fleets, ,
  drop = FALSE
]
if (nrow(ct_comm) > 0) {
  mv_plot_trend_multi(
    df      = ct_comm,
    group   = "fleet",
    ylab    = "Catch (t)",
    colours = mv_fleet_colours(
      fleet_names = comm_fleets
    )
  )
}
```

## Row {eval=has_tac}

### TAC

```{r tac_envelope, eval=has_tac}
p <- mv_plot_trend(
  df     = mv$tac,
  ylab   = "TAC (t)",
  colour = "#9467bd"
)
p <- plotly_build(p)
p$x$layout$shapes <- c(p$x$layout$shapes, list(tMP_vline))
p
```


# Indices {`r if (!has_indices) ".hidden"`}

```{r index_helpers, eval=has_indices, include=FALSE}
idx_cols <- mv_fleet_colours(
  fleet_names = survey_fleets
)
has_survey_data <- !is.null(mv$survey_data) &&
  length(mv$survey_data) > 0

# Generic index slider from survey_data bundle
.idx_slider <- function(fl, col) {
  sd <- mv$survey_data[[fl]]
  if (is.null(sd)) return(NULL)

  tr_state <- sd$state
  tr_idx   <- sd$index
  tr_q     <- sd$q
  unit     <- sd$label

  y_label <- paste0(
    fl, " (", unit, ")"
  )

  p <- plot_ly()
  for (i in seq_along(good_reps)) {
    r <- good_reps[i]
    st_sub <- tr_state[tr_state$rep == r, , drop = FALSE]
    ix_sub <- tr_idx[tr_idx$rep == r, , drop = FALSE]
    q_sub  <- tr_q[tr_q$rep == r, , drop = FALSE]
    scaled <- ix_sub$value / q_sub$value
    iv <- which(!is.na(scaled) & scaled > 0)

    p <- add_lines(
      p = p,
      x = st_sub$year, y = st_sub$value,
      visible = (i == 1),
      name = paste0("Vuln. ", unit),
      line = list(color = col, width = 2),
      showlegend = FALSE
    )
    if (length(iv) > 0) {
      p <- add_markers(
        p = p,
        x = ix_sub$year[iv], y = scaled[iv],
        visible = (i == 1),
        name = "Index / q",
        marker = list(
          color = col, size = 5,
          symbol = "circle-open",
          line = list(width = 1.5, color = col)
        ),
        showlegend = FALSE
      )
    } else {
      # Placeholder trace (invisible) to keep trace count
      p <- add_markers(
        p = p,
        x = NA, y = NA,
        visible = FALSE,
        name = "Index / q",
        showlegend = FALSE
      )
    }
  }

  # Slider: toggle pairs of traces (vB + idx per rep)
  n_per_rep <- 2L
  idx_steps <- lapply(
    X = seq_along(good_reps),
    FUN = function(i) {
      vis <- rep(FALSE, length(good_reps) * n_per_rep)
      vis[((i - 1) * n_per_rep + 1):(i * n_per_rep)] <-
        TRUE
      list(
        method = "restyle",
        args = list(list(visible = as.list(vis))),
        label = as.character(good_reps[i])
      )
    }
  )

  lay_i <- mv_layout()
  lay_i$xaxis$title <- "Year"
  lay_i$yaxis$title <- y_label
  lay_i$xaxis$range <- list(min(years), max(years))
  lay_i$showlegend <- FALSE
  lay_i$shapes <- list(tMP_vline)
  lay_i$sliders <- list(list(
    active = 0,
    currentvalue = list(
      prefix = "Replicate: ",
      font = list(size = 12)
    ),
    pad = list(t = 40),
    steps = idx_steps
  ))
  p <- do.call(
    what = layout,
    args = c(list(p = p), lay_i)
  )
  mv_config(p = p)
}
```

## Row

### `r survey_fleets[1]` survey

```{r idx_panel_1, eval=has_indices && has_survey_data && length(survey_fleets) >= 1}
.idx_slider(fl = survey_fleets[1], col = idx_cols[1])
```

## Row

### `r if (length(survey_fleets) >= 2) survey_fleets[2] else ""`

```{r idx_panel_2, eval=has_indices && has_survey_data && length(survey_fleets) >= 2}
.idx_slider(fl = survey_fleets[2], col = idx_cols[2])
```


# Example replicate

## Row {data-height=50}

###

Use the slider below each plot to page through
replicates.

## Row

### Spawning biomass

```{r rep_sb_slider}
sb_traces <- mv$traces$spawning_biomass
p <- plot_ly()
for (i in seq_along(good_reps)) {
  r <- good_reps[i]
  sub <- sb_traces[sb_traces$rep == r, , drop = FALSE]
  p <- add_lines(
    p = p,
    x = sub$year, y = sub$value,
    visible = (i == 1),
    name = paste0("Rep ", r),
    line = list(color = "#1f77b4", width = 2),
    hovertemplate = paste0(
      "Rep ", r, " %{x}: %{y:.3g}<extra></extra>"
    )
  )
}

steps <- lapply(
  X = seq_along(good_reps),
  FUN = function(i) {
    vis <- rep(FALSE, length(good_reps))
    vis[i] <- TRUE
    list(
      method = "restyle",
      args = list(list(visible = as.list(vis))),
      label = as.character(good_reps[i])
    )
  }
)

lay <- mv_layout()
lay$xaxis$title <- "Year"
lay$yaxis$title <- "Spawning biomass (t)"
lay$showlegend <- FALSE
lay$shapes <- list(tMP_vline)
lay$sliders <- list(list(
  active = 0,
  currentvalue = list(
    prefix = "Replicate: ",
    font = list(size = 12)
  ),
  pad = list(t = 40),
  steps = steps
))
if (!is.null(Bmsy)) {
  lay$shapes <- c(lay$shapes, list(list(
    type = "line",
    x0 = min(years), x1 = max(years),
    y0 = Bmsy, y1 = Bmsy,
    line = list(
      color = "#999999", width = 1.5, dash = "dash"
    )
  )))
}

p <- do.call(
  what = layout,
  args = c(list(p = p), lay)
)
mv_config(p = p)
```

### Catch

```{r rep_catch_slider}
ct_traces <- mv$traces$catch
p <- plot_ly()
for (i in seq_along(good_reps)) {
  r <- good_reps[i]
  sub <- ct_traces[ct_traces$rep == r, , drop = FALSE]
  p <- add_lines(
    p = p,
    x = sub$year, y = sub$value,
    visible = (i == 1),
    name = paste0("Rep ", r),
    line = list(color = "#d62728", width = 2),
    hovertemplate = paste0(
      "Rep ", r, " %{x}: %{y:.3g}<extra></extra>"
    )
  )
}

lay2 <- mv_layout()
lay2$xaxis$title <- "Year"
lay2$yaxis$title <- "Catch (t)"
lay2$showlegend <- FALSE
lay2$shapes <- list(tMP_vline)
lay2$sliders <- list(list(
  active = 0,
  currentvalue = list(
    prefix = "Replicate: ",
    font = list(size = 12)
  ),
  pad = list(t = 40),
  steps = steps
))

p <- do.call(
  what = layout,
  args = c(list(p = p), lay2)
)
mv_config(p = p)
```

## Row

### Recruitment

```{r rep_rec_slider}
rc_traces <- mv$traces$recruitment
p <- plot_ly()
for (i in seq_along(good_reps)) {
  r <- good_reps[i]
  sub <- rc_traces[rc_traces$rep == r, , drop = FALSE]
  p <- add_lines(
    p = p,
    x = sub$year, y = sub$value,
    visible = (i == 1),
    name = paste0("Rep ", r),
    line = list(color = "#2ca02c", width = 2),
    hovertemplate = paste0(
      "Rep ", r, " %{x}: %{y:.3g}<extra></extra>"
    )
  )
}

lay3 <- mv_layout()
lay3$xaxis$title <- "Year"
lay3$yaxis$title <- "Recruitment (age 0)"
lay3$showlegend <- FALSE
lay3$shapes <- list(tMP_vline)
lay3$sliders <- list(list(
  active = 0,
  currentvalue = list(
    prefix = "Replicate: ",
    font = list(size = 12)
  ),
  pad = list(t = 40),
  steps = steps
))

p <- do.call(
  what = layout,
  args = c(list(p = p), lay3)
)
mv_config(p = p)
```

### Stock-recruitment

```{r rep_sr_slider, eval=has_sr}
sr_data <- mv$stock_recruit
p <- plot_ly()

# BH curve (always visible, first trace)
has_bh <- has_refpts &&
  !is.na(mv$ref_points$rec_a) &&
  !is.na(mv$ref_points$rec_b)
if (has_bh) {
  reca <- mv$ref_points$rec_a
  recb <- mv$ref_points$rec_b
  ssb_seq <- seq(
    from = 0,
    to   = max(sr_data$ssb, na.rm = TRUE) * 1.1,
    length.out = 200
  )
  r_pred <- reca * ssb_seq / (1 + recb * ssb_seq)
  p <- add_lines(
    p = p,
    x = ssb_seq, y = r_pred,
    name = "BH curve",
    line = list(color = "#d62728", width = 2),
    hovertemplate = paste0(
      "BH: SSB=%{x:.3g} R=%{y:.3g}<extra></extra>"
    )
  )
}

# Replicate scatter traces
for (i in seq_along(good_reps)) {
  r <- good_reps[i]
  sub <- sr_data[sr_data$rep == r, , drop = FALSE]
  p <- add_markers(
    p = p,
    x = sub$ssb, y = sub$recruitment,
    visible = (i == 1),
    name = paste0("Rep ", r),
    color = I(ifelse(
      test = sub$period == "Historical",
      yes  = "#1f77b4", no = "#999999"
    )),
    marker = list(size = 5),
    text = sub$year,
    hovertemplate = paste0(
      "Rep ", r,
      "<br>Year: %{text}",
      "<br>SSB: %{x:.3g}",
      "<br>R: %{y:.3g}<extra></extra>"
    )
  )
}

# Slider steps: BH curve always visible (trace 0),
# rep traces toggled (traces 1..nReps)
sr_steps <- lapply(
  X = seq_along(good_reps),
  FUN = function(i) {
    vis <- rep(FALSE, length(good_reps))
    vis[i] <- TRUE
    # Prepend TRUE for BH curve if present
    if (has_bh) vis <- c(TRUE, vis)
    list(
      method = "restyle",
      args = list(list(visible = as.list(vis))),
      label = as.character(good_reps[i])
    )
  }
)

lay4 <- mv_layout()
lay4$xaxis$title <- "Spawning biomass (t)"
lay4$yaxis$title <- "Recruitment (age 0)"
lay4$showlegend <- FALSE
lay4$sliders <- list(list(
  active = 0,
  currentvalue = list(
    prefix = "Replicate: ",
    font = list(size = 12)
  ),
  pad = list(t = 40),
  steps = sr_steps
))

p <- do.call(
  what = layout,
  args = c(list(p = p), lay4)
)
mv_config(p = p)
```


# Conditioning {`r if (!has_cond_valid) ".hidden"`}

## Row {data-height=60}

###

Historical MS3 operating model states vs NCAM assessment
model estimates (replicate 1). Relative error =
(MS3 - NCAM) / |NCAM|.

```{r re_range, eval=has_cond_valid, include=FALSE}
# Per-state RE axis range: [-1, 1] or wider if needed
cv <- mv$cond_valid
.re_lim <- function(re) {
  max(1, ceiling(max(abs(re), na.rm = TRUE) * 10) / 10)
}
re_lim_ssb <- .re_lim(re = cv$re_ssb)
re_lim_R   <- .re_lim(re = cv$re_R)
re_lim_N   <- .re_lim(re = cv$re_N)
re_lim_C   <- if (!is.null(cv$re_C)) .re_lim(re = cv$re_C) else 1
```

## Row

### Spawning biomass: MS3 vs NCAM

```{r cond_ssb, eval=has_cond_valid}
cv <- mv$cond_valid
p <- plot_ly() %>%
  add_lines(
    x = cv$years, y = cv$ncam_ssb,
    name = "NCAM",
    line = list(color = "#d62728", width = 2)
  ) %>%
  add_lines(
    x = cv$years, y = cv$ms3_ssb,
    name = "MS3",
    line = list(color = "#1f77b4", width = 2,
                dash = "dash")
  )
lay <- mv_layout()
lay$xaxis$title <- "Year"
lay$yaxis$title <- "Spawning biomass (t)"
lay$showlegend <- TRUE
p <- do.call(what = layout, args = c(list(p = p), lay))
mv_config(p = p)
```

### SSB relative error

```{r cond_ssb_re, eval=has_cond_valid}
cv <- mv$cond_valid
df <- data.frame(year = cv$years, est = cv$re_ssb)
p <- mv_plot_trend(
  df = df, ylab = "Relative error",
  colour = "#333333", ref_line = 0
)
p <- layout(p, yaxis = list(range = list(-re_lim_ssb, re_lim_ssb)))
p
```

## Row

### Recruitment: MS3 vs NCAM

```{r cond_rec, eval=has_cond_valid}
cv <- mv$cond_valid
p <- plot_ly() %>%
  add_lines(
    x = cv$years, y = cv$ncam_R,
    name = "NCAM",
    line = list(color = "#d62728", width = 2)
  ) %>%
  add_lines(
    x = cv$years, y = cv$ms3_R,
    name = "MS3",
    line = list(color = "#2ca02c", width = 2,
                dash = "dash")
  )
lay <- mv_layout()
lay$xaxis$title <- "Year"
lay$yaxis$title <- "Recruitment (age 0)"
lay$showlegend <- TRUE
p <- do.call(what = layout, args = c(list(p = p), lay))
mv_config(p = p)
```

### Recruitment relative error

```{r cond_rec_re, eval=has_cond_valid}
cv <- mv$cond_valid
df <- data.frame(year = cv$years, est = cv$re_R)
p <- mv_plot_trend(
  df = df, ylab = "Relative error",
  colour = "#333333", ref_line = 0
)
p <- layout(p, yaxis = list(range = list(-re_lim_R, re_lim_R)))
p
```

## Row

### Total N: MS3 vs NCAM

```{r cond_n, eval=has_cond_valid}
cv <- mv$cond_valid
p <- plot_ly() %>%
  add_lines(
    x = cv$years, y = cv$ncam_N,
    name = "NCAM",
    line = list(color = "#d62728", width = 2)
  ) %>%
  add_lines(
    x = cv$years, y = cv$ms3_N,
    name = "MS3",
    line = list(color = "#ff7f0e", width = 2,
                dash = "dash")
  )
lay <- mv_layout()
lay$xaxis$title <- "Year"
lay$yaxis$title <- "Total numbers"
lay$showlegend <- TRUE
p <- do.call(what = layout, args = c(list(p = p), lay))
mv_config(p = p)
```

### Total N relative error

```{r cond_n_re, eval=has_cond_valid}
cv <- mv$cond_valid
df <- data.frame(year = cv$years, est = cv$re_N)
p <- mv_plot_trend(
  df = df, ylab = "Relative error",
  colour = "#333333", ref_line = 0
)
p <- layout(p, yaxis = list(range = list(-re_lim_N, re_lim_N)))
p
```

## Row

### Catch: MS3 vs NCAM

```{r cond_catch, eval=has_cond_valid && !is.null(mv$cond_valid$ncam_C)}
cv <- mv$cond_valid
p <- plot_ly() %>%
  add_lines(
    x = cv$years, y = cv$ncam_C,
    name = "NCAM",
    line = list(color = "#d62728", width = 2)
  ) %>%
  add_lines(
    x = cv$years, y = cv$ms3_C,
    name = "MS3",
    line = list(color = "#9467bd", width = 2,
                dash = "dash")
  )
lay <- mv_layout()
lay$xaxis$title <- "Year"
lay$yaxis$title <- "Catch (t)"
lay$showlegend <- TRUE
p <- do.call(what = layout, args = c(list(p = p), lay))
mv_config(p = p)
```

### Catch relative error

```{r cond_catch_re, eval=has_cond_valid && !is.null(mv$cond_valid$re_C)}
cv <- mv$cond_valid
df <- data.frame(year = cv$years, est = cv$re_C)
p <- mv_plot_trend(
  df = df, ylab = "Relative error",
  colour = "#333333", ref_line = 0
)
p <- layout(p, yaxis = list(range = list(-re_lim_C, re_lim_C)))
p
```


# Performance {`r if (pT < 15) ".hidden"`}

## Row {data-height=80}

###

Performance summary over `r nReps` replicates across
the `r pT`-year projection period
(`r tMP_year`--`r max(years)`).

## Row

### Spawning biomass relative to Bmsy

```{r depletion, eval=has_refpts}
dep <- mv$spawning_biomass
dep$est <- dep$est / Bmsy
dep$lwr <- dep$lwr / Bmsy
dep$upr <- dep$upr / Bmsy
p <- mv_plot_trend(
  df       = dep,
  ylab     = "SB / Bmsy",
  ref_line = 1.0,
  ref_label = "Bmsy"
)
lrp_line <- list(
  type = "line",
  x0 = min(years), x1 = max(years),
  y0 = 0.4, y1 = 0.4,
  line = list(
    color = "#d62728", width = 1, dash = "dash"
  )
)
p <- layout(p, shapes = list(tMP_vline, lrp_line))
p
```

### Catch variability

```{r catch_var, eval=has_catch}
proj_idx <- which(years >= tMP_year)
if (length(proj_idx) > 1) {
  ct_proj <- mv$catch[proj_idx, , drop = FALSE]
  ct_proj$cv <- (ct_proj$upr - ct_proj$lwr) /
    (2 * ct_proj$est + 1e-10)
  cv_df <- data.frame(
    year = ct_proj$year,
    est  = ct_proj$cv
  )
  mv_plot_trend(
    df     = cv_df,
    ylab   = "Catch envelope width (relative)",
    colour = "#8c564b"
  )
}
```
